---
title: "Assignment 9"
author: "Thomas Caldwell"
date: "4/24/2020"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(modelr)
library(lindia)
library(MASS)
df = read.csv("./Data/GradSchool_Admissions.csv")
df$admit = as.logical(df$admit)
```

### Dataset

The dataset we are working with is Graduate School admission data.

```{r Dataframe, echo=TRUE, warning=FALSE}
colnames(df)
```

As we can see the data includes wether or not they were admited, their gre score, their gpa, and what rank their school was.

I want to see which specific one is a bigger indicator and then find a model that can most accuratley predict acceptance based on all the factors.

### Exploratory plots

To first see any correlations I'm going to plot the three factors individually to acceptance.

```{r plots, echo=FALSE}
plot(df$gre, df$admit)
abline(lm(df$admit~df$gre))
plot(df$gpa, df$admit)
abline(lm(df$admit~df$gpa))
plot(df$rank, df$admit)
abline(lm(df$admit~df$rank))
```

From this we can tell that gre an gpa have a positive correlation and rank and a negative but that's because the ranks are highest to lowest.

### Models

All the factors seemed relevant, but to find the best model we will use stepAIC. We will first make a model including all the factors and then run it through to determine the best.

```{r model, echo=TRUE, message=FALSE, warning=FALSE}
mod = lm(admit ~ gre * gpa * rank ,df)
step = stepAIC(mod, direction="both")
```

The selected model:

```{r model anova, echo=TRUE, message=TRUE, warning=FALSE}
step$call
step$anova
```

### Testing the model

### Conclusion








