---
title: "Final"
output:
  html_document:
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
theme_set(theme_minimal())

dfn = readRDS("../Final/RDS_data/cleaned_nic_data.RDS")
dfo = readRDS("../Final/RDS_data/cleaned_odor_data.RDS")
NMDS = readRDS("../Final/RDS_data/nic_MDS.RDS")
OMDS = readRDS("../Final/RDS_data/odor_MDS.RDS")
nadonis = readRDS("../Final/RDS_data/nic_anova.RDS")
oadonis = readRDS("../Final/RDS_data/odor_anova.RDS")


```

[Cleaning the data](#cleaningthedata)

[Processing the data](#processingthedata)

[Results](#results)

[Conclusion](#conclusion)

This project is analyzing data obtained from Dr Dani Taylor. The expiriment is testing how a bee brain reacts to nicotine and a specific odor after the bee has been admisnistered octopamine.

The following graphs are an overview of the data.

```{r echo=FALSE}
ggplot(dfn, aes(x=Time, y=nicotine, color=prepost))+
  geom_line()+
  facet_wrap(~Bee)

ggplot(dfo, aes(x=Time, y=odor, color=prepost))+
  geom_line()+
  facet_wrap(~Bee)
```

## Cleaning the data {#cleaningthedata}

When I received the data it was in .pzfz format. To read it to r I used the pzfx package.

```{r eval=FALSE}
library(pzfx)
```

It was in multiple files in multiple tables so I used the following code to upload it.

```{r eval=FALSE}
#i is used to assign a number to each bee
i=0

#read in first dataset
tabs = pzfx_tables("./data/NON_11-08-11.pzfx")
for(Table in tabs){
  i=i+1
  if(Table == tabs[1]){
    df = cbind(Bee=i, Date="11-08-11", Table, 
               read_pzfx("./data/NON_11-08-11.pzfx", table=Table))
  }
  else{
    df = rbind(df, cbind(Bee=i, Date="11-08-11", Table, 
               read_pzfx("./data/NON_11-08-11.pzfx", table=Table)))
  }
  
}

#read in second dataset
tabs = pzfx_tables("./data/NON_11-22-11.pzfx")
for(Table in tabs){
  i=i+1
  df = rbind(df, cbind(Bee=i, Date="11-22-11", Table, 
                       read_pzfx("./data/NON_11-22-11.pzfx", table=Table)))
}

#read in third dataset
tabs = pzfx_tables("./data/NON_1-11-12.pzfx")
for(Table in tabs){
  i=i+1
  df = rbind(df, cbind(Bee=i, Date="1-11-12", Table, 
                       read_pzfx("./data/NON_1-11-12.pzfx", table=Table)))
}

rm(tabs, Table, i)
```

After I succesfully read in the data made some corrections and cleaned up the data as follows

```{r eval=FALSE}
#Time in seconds
names(df)[4] = "Time"
names(df)[16] = "o3"
names(df)[8] = "nicotine"
names(df)[21] = "nicotine"
names(df)[12] = "odor"
names(df)[17] = "odor"

#fix averages
df[8] = rowMeans(df[5:7])
df[12] = rowMeans(df[9:11])
df[17] = rowMeans(df[14:16])
df[21] = rowMeans(df[18:20])

#prepost nicodor mod
dfn = rbind(cbind(df[,-c(9:12,14:17,18:21)], prepost="pre"), cbind(df[,-c(5:8,9:12,14:17)], prepost="post"))
dfo = rbind(cbind(df[,-c(5:8,14:17,18:21)], prepost="pre"), cbind(df[,-c(5:8,9:12,18:21)], prepost="post"))

dfn = dfn[order(dfn$Bee),]
dfo = dfo[order(dfo$Bee),]
dfn = dfn[complete.cases(dfn),]
dfo = dfo[complete.cases(dfo),]
```

## Processing the data {#processingthedata}

To analyze this data I used the vegan pacakge.

```{r}
library(vegan)
```

I used both metaMDS and adonis to determine the relationship between how the bee brain reacted prior to the octopamine and after.

```{r eval=FALSE}
NMDS = metaMDS(dfn[5:7],k = 2,distance = "euclidean",na.rm = TRUE) #k=3 third axis
OMDS = metaMDS(dfo[5:7],k = 2,distance = "euclidean",na.rm = TRUE)

ndist = vegdist(dfn[5:7], method = "euclidean")
odist = vegdist(dfo[5:7], method = "euclidean")
nadonis = adonis(ndist ~ dfn$prepost)
oadonis = adonis(odist ~ dfo$prepost)
```

## Results {#results}

First let's review the results of the permanova we obtained through adonis.

```{r}
nadonis
oadonis
```

As we can see the permanova showed that the octopamine administration could only show 1-2 percent of the varience. The following is the data from the metaMDS. First the stressplots from the nicotine dataset then the odor dataset.

```{r}
stressplot(NMDS)
stressplot(OMDS)
```

Now we will plot the two points from the MDS.

```{r message=FALSE, warning=FALSE}
data.frame(MDS_Point_1=NMDS$points[,1],MDS_Point_2=NMDS$points[,2],PrePost = dfn$prepost) %>%
ggplot(aes(x=MDS_Point_1,y=MDS_Point_2,color=PrePost))+
  geom_point()+
  stat_ellipse()+
  ggtitle("Nicotine")

data.frame(MDS_Point_1=OMDS$points[,1],MDS_Point_2=OMDS$points[,2],PrePost = dfo$prepost) %>%
ggplot(aes(x=MDS_Point_1,y=MDS_Point_2,color=PrePost))+
  geom_point()+
  stat_ellipse()+
  ggtitle("Odor")+
  xlim(-0.2, 0.2)+
  ylim(-0.2, 0.2)
```

There is a lot of overlap but the permanova showed that there was a difference.

## Conclusion {#conclusion}

The permanova was able to show that the octopamine had an impact on the results but with a R^2 of 1-2 percent it isn't nearly as impactful as the graphs at the top show. I plan on using another package such as ARIMA to analyze the data. ARIMA will be nice because  we are processing a time series.

Thanks again to Dr Dani Taylor for providing the data.
